Sparse Word Embeddings for Multi-domain Machine Translation

### 2019-05-06 2019/emnlp/sparse.tex
new

\section{Introduction \label{sec:introduction}}
% The problem
Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date. One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain. A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{Firat16multiway,Johnson17google} or even building systems without any data \cite{Artetxe18unsupervised,Lample18unsupervised}.

Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains. Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT). The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain. We study here a different scenario, multi-domain adaptation \cite{}Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains. This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain). In our setting, we assume that the domain of test documents is known in advance.

@@

\begin{algorithm}[H]
\caption{Multi-domain Training}
\label{algo:1}
\begin{algorithmic}[1]
\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for d domains}
\REPEAT 
\STATE{Randomly pick $i \in [1,..,d]$ w.r.t multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
\STATE{Randomly pick a batch of sentences from corpora $C_i$.}
\STATE{Activating only generic activate embedding to create generic embedding batch, denoted $W_g$}
\STATE{Pass the batch of word embedding to a dense layer to fuse regions}
\STATE{Computing gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using generic batch $W_g$}
\STATE{Activating corresponding domain region and generic region to create domain embedding batch}
\STATE{Pass the batch of word embedding to a dense layer to fuse regions.}
\STATE{Computing gradient of domain parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using domain batch }
\STATE{Apply backprogation using computed gradients $\frac{\partial L}{\partial \theta_s}$ and $\frac{\partial L}{\partial \theta_i}$}
\UNTIL{convergence}
\end{algorithmic}
\end{algorithm}

@@

\subsection{Results}
\begin{table}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
 \hline
 ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
 \hline
 Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by a pre-trained domain classififer}
\label{tab:1}
\end{table}

@@

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 65.42 & 34.7 & 51.38 & 29.27 & 28.95 & 17.82 & 22.21 & 21.63 \\
 \hline
 EMEA-finetuned & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse & 68.51 & 35.75 & 51.19 & 28.88 & 28.92 & 17.89 & 22.73 & 21.08\\
 \hline
 Domain Control & & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
 \hline 
\end{tabularx}}
\end{center}
\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French. Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{Zeng18multidomain} use Bidirectional-LSTM}
\label{tab:3}
\end{figure}

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 62.63 & 23.91 & 66.95 & 16.92 & 21.29 & 10.9 & & 14.19 \\
 \hline
 EMEA-finetuned & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse & & & & & & & & \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & & & & & & & & \\
 \hline 
\end{tabularx}}
\end{center}
\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
\label{tab:4}
\end{figure}

>>>>

\section{Introduction \label{sec:introduction}}
% The problem
Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date. One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain. A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{Firat16multiway,Johnson17google} or even building systems without any data \cite{Artetxe18unsupervised,Lample18unsupervised}.
Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains. Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT). The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain. We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains. This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain). In our setting, we assume that the domain of test documents is known in advance.

@@

\begin{algorithm}[h]
\caption{Multi-domain Training}
\label{algo:1}
\begin{algorithmic}[1]
\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for d domains}
\REPEAT 
\STATE{Randomly pick $i \in [1,..,d]$ w.r.t multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
\STATE{Randomly pick a batch of sentences from corpora $C_i$.}
\STATE{Activating only generic activate embedding to create generic embedding batch, denoted $W_g$}
\STATE{Pass the batch of word embedding to a dense layer to fuse regions}
\STATE{Computing gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using generic batch $W_g$}
\STATE{Activating corresponding domain region and generic region to create domain embedding batch}
\STATE{Pass the batch of word embedding to a dense layer to fuse regions.}
\STATE{Computing gradient of domain parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using domain batch }
\STATE{Apply backprogation using computed gradients $\frac{\partial L}{\partial \theta_s}$ and $\frac{\partial L}{\partial \theta_i}$}
\UNTIL{convergence}
\end{algorithmic}
\end{algorithm}

@@

\subsection{Results}
\begin{table}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.10 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
 \hline
 ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
 \hline
 Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by a pre-trained domain classififer}
\label{tab:1}
\end{table}

@@

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 65.42 & 34.7 & 51.38 & 29.27 & 28.95 & 17.82 & 22.21 & 21.63 \\
 \hline
 EMEA-finetuned  & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse (1) & 68.51 & 35.75 & 51.19 & 28.88 & 28.92 & 17.89 & 22.73 & 21.08 \\
 \hline
 Sparse (2) & & & & & & & & \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
 \hline 
\end{tabularx}}
\end{center}
\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French. Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{Zeng18multidomain} use Bidirectional-LSTM}
\label{tab:3}
\end{figure}

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 57.37 & 23.1 & 63.54 & 17.39 & 20.65 & 10.69 & 11.51 & 14.29 \\
 \hline
 EMEA-finetuned & 65.64 & 12.36 & 15.93 & 15.38 & 11.1 & 5.54 & 6.22 & 7.6 \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse (1) & 63.43 & 22.66 & 64.4 & 16.99 & 20.63 & 10.89 & 11.45 & 14.26 \\
 \hline
 Sparse (2) & 63.27 & 22.62 & 63.7 & 17.26 & 20.69 & 10.89 & 11.45 & 14.26 \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & & & & & & & & \\
 \hline 
\end{tabularx}}
\end{center}
\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
\label{tab:4}
\end{figure}
